---
title: "Eric_Hirsch_624_Homework_5"
output: html_document
date: "2023-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}

library(tidyverse)
library(tsibble)
library(fpp3)
library(EHData)
library(gridExtra)
library(fable)
library(readxl)

```

First we read in the data and look for missing values.  

There are 19 records with missing values in the dataset, including 14 with a missing ATM (they all represent days from the month of May 2010, which is what we are predicting.)  This leaves 5 missing cash values - we use ARIMA to impute them.

Then we separate the original tsibble into 4 tsibbles, each representing one ATM.

```{r}

dfATM <- as.data.frame(read_excel("D:\\RStudio\\CUNY_624\\Project1\\ATM624Data.xlsx"))

summary(dfATM)
str(dfATM)

print("Missing ATM: ")
which(is.na(dfATM$ATM))
print("Missing Cash: ")
which(is.na(dfATM$Cash))

CreateAtmTsibble <- function(filter) {
  
  ts <- dfATM %>%
  filter(ATM==filter)%>%
  mutate(xDate=as_date(DATE)) %>%
  as_tsibble(index=xDate) %>%
  dplyr::select(xDate, Cash)
  
  ts <- ts %>%
  model(ARIMA(Cash)) %>%
  interpolate(ts)
  
  return (ts)
}

tsATM1a <- CreateAtmTsibble("ATM1")
tsATM2a <- CreateAtmTsibble("ATM2")
tsATM3 <- CreateAtmTsibble("ATM3")
tsATM4 <- CreateAtmTsibble("ATM4")


```


## Forecasts {.tabset}

### ATM1

1. Examine the data.

First we look a the data. It shows a number of strange patterns, not the least of which is that the money removed from the ATMs is not in multiples of twenty.  Also, the distribution is bimodal.  There are possible outliers but none of them are too out of range. Since we have no contextual information, we simply move forward.

We will train the model on 75% of the data.

```{r, fig.height=2}

Summarize_ATM <- function(ts, ATM) {
dfCash1 <- as.data.frame(ts$Cash)
a <- EHSummarize_SingleColumn_Boxplots(dfCash1)
b <- EHSummarize_SingleColumn_Histograms(dfCash1)

grid.arrange(grobs=c(a,b), ncol=2)

train1 <- ts %>%
  filter_index(~ "2010-01-31")

test1 <- ts %>%
  filter_index("2010-02-1" ~ .)

set.seed(042760)


  stl<- tsATM1a %>%
  stl(s.window ="periodic")

c <- forecast::autoplot(ts, Cash) +
  autolayer(test1, Cash, colour = "red") + labs(title = paste(ATM, ", Training Data is <= January")) 
print(c)

x <- list(train1, test1, stl)
return(x)
}


```

2. Now we try different models.

```{r}

Train_ATM <- function(train1){
fit <- train1 %>%
  model(SNAIVE = SNAIVE(), ETS_AAN = ETS(Cash ~ error("A") + trend("A") + season("N")), ETS_AAA = ETS(Cash ~ error("A") + trend("A") + season("A")), ETS_AAdA = ETS(Cash ~ error("A") + trend("Ad", phi=.85) + season("A")), ARIMA=ARIMA(Cash))

return(fit)
}

```

```{r, include=FALSE}

Examine_Residuals <- function(fit) {

resids <- function(model) {
a <- fit[model] %>% gg_tsresiduals + labs(title=model)
return (a)
}

z <- list()
plot_list <- list()

for (i in names(fit)) {
  index=1
  
  x <- resids(i)
  p <- eval(substitute(x, list(i=index)))
  plot_list[[i]] <- p 
  
  index=index+1
}
return (plot_list)
}


```

```{r, fig.height=9}
Print_Residuals <- function(plot_list){
grid.arrange(grobs=c(plot_list$SNAIVE[1:3], plot_list$ETS_AAN[1:3], plot_list$ETS_AAA[1:3], plot_list$ETS_AAdA[1:3], plot_list$ARIMA[1:3]), ncol=3)
}
```

```{r, fig.height=4}
Check_RMSE <- function(fit, test1, ts, ATM){
fc <- fit %>%
  forecast(test1)

a <- fc %>% forecast::autoplot(ts) + labs(title = paste(ATM, ", Training Data is <= January"))

b <- fit %>% accuracy()
c <- fc %>% accuracy(test1)

ww <- list(a,b,c)

}
```


```{r, fig.height=2}
train <- Summarize_ATM(tsATM1a, "ATM1")
```

```{r, fig.height = 2}

autoplot(train[[3]])
```

```{r}
fit <- Train_ATM(train[[1]])
```
```{r, include=FALSE}
pp <- Examine_Residuals(fit)
```
```{r, fig.height=9}
Print_Residuals(pp)
```
```{r, fig.height=4}
ww <- Check_RMSE(fit, train[[2]], tsATM1a, "ATM1")
plot(ww[[1]])
ww[[3]]
```
3. Forecast using the best model.

Interestingly, the RMSEs are comparable for all models except ETS_AAN (even SNAIVE).  However, the residuals indicate a strong issue (obviously no accounting for seasonality) so we don't trust standard errors and other statistical inference calculations.

In this case we might compare the results of several modes before deciding what to do.  First we look at total May cash using ARIMA(0-0-3,0-1-1)(7)).  The sum is $2,402.

```{r}

fit4d <- tsATM1a %>%
  model(ARIMA(Cash ~ pdq(0,0,3) + PDQ(0,1,1)))
  
fc4d <- fit4d %>%
  forecast(h = 31)

fc4d %>%
  forecast::autoplot(tsATM1a) +
  geom_line(aes(y = .fitted), col="#D55E00",
            data = augment(fit4d)) +
  labs(y="Count", title="ATM1 - ARIMA(0-0-3, 0-1-1)(7)") +
  guides(colour = "none")

report(fit4d)
sum(fc4d$.mean)

```
Now we look at total May cash using ETS(AAN).  The sum is $2,465.  There is only a $63 difference despite the lack of a seasonal component in the ATM. We can see from the Monthly cash plot that May's cash haul last year was comparable, so we split the difference - $2,435.

```{r}

fit4d <- tsATM1a %>%
  model(ETS(Cash ~ error("A") + trend("A") + season("N")))
  
fc4d <- fit4d %>%
  forecast(h = 31)

fc4d %>%
  forecast::autoplot(tsATM1a) +
  geom_line(aes(y = .fitted), col="#D55E00",
            data = augment(fit4d)) +
  labs(y="Count", title="ATM1 - ETS-AAN") +
  guides(colour = "none")

report(fit4d)
sum(fc4d$.mean)

```

```{r}

tsATM1b <- tsATM1a %>%
  as.data.frame() %>%
  group_by(Month=yearmonth(xDate)) %>%
  dplyr::summarize(MonthlyCash=sum(Cash)) %>%
  as_tsibble(index="Month")

forecast::autoplot(tsATM1b)

```


### ATM2

1. Examine the data.

First we look a the data. It shows similar irregularities as ATM1, including the bimodal distribution and cash not in multiples of twenty.  Also, the distribution is bimodal.  There are no apparent outliers.  Variance may be falling over time so a boxcox is employed here.

We will train the model on 75% of the data.

```{r, fig.height=2}
train <- Summarize_ATM(tsATM2a, "ATM2")
```
```{r}
lambda <- tsATM2a %>%
  features(Cash, features = guerrero) %>%
  pull(lambda_guerrero)

tsATM2b <- tsATM2a %>%
 mutate(Cash=box_cox(Cash, lambda))

train <- Summarize_ATM(tsATM2b, "ATM2 with BoxCox")

fit <- Train_ATM(train[[1]])
```
```{r, include=FALSE}
pp <- Examine_Residuals(fit)
```
```{r, fig.height=9}
Print_Residuals(pp)
```
```{r, fig.height=4}
ww <- Check_RMSE(fit, train[[2]], tsATM2a, "ATM2")
plot(ww[[1]])
ww[[3]]
```
3. Forecast using the best model.

The RMSEs are again comparable for all models - ETS_AAN again performs best but the residuals suggest the standard error may not be accurate. 

We will estimate the cash using ETS_AADA as it had the second lowest RMSE and the residuals indicate white noise.  We will need to reverse the boxcox.

Total cash for May is $1,825. This falls short of last May but reflects a significant downward trend after May.

```{r}
fit2 <- tsATM2b %>%
  model(ETS(Cash ~ error("A") + trend("Ad", phi=.85) + season("A")))

fc2 <- fit2 %>%
  forecast(h = 31)

rr <- as.vector(fc2$.mean)
rr2 <- forecast::InvBoxCox(rr, lambda)

fc2 %>%
  forecast::autoplot(tsATM2b) +
  geom_line(aes(y = .fitted), col="#D55E00",
            data = augment(fit2)) +
  labs(y="Count", title="ATM2 - ETS(AAdA), phi=.85") +
  guides(colour = "none")

report(fit2)
sum(rr2)

```


```{r}

tsATM2c <- tsATM2a %>%
  as.data.frame() %>%
  group_by(Month=yearmonth(xDate)) %>%
  dplyr::summarize(MonthlyCash=sum(Cash)) %>%
  as_tsibble(index="Month")

forecast::autoplot(tsATM2c)

```


### ATM3

1. Examine the data.

ATM3 has three values for April and nothing else. It is tempting to simply take the mean of those three values and apply them to May. While that might suffice for April's estimate, I am going to go out on a limb and suggest that 

```{r, fig.height=2}
train <- Summarize_ATM(tsATM3, "ATM3")
```

```{r}


tsATM4c <- tsATM4 %>%
  as.data.frame() %>%
  group_by(Month=yearmonth(xDate)) %>%
  dplyr::summarize(MonthlyCash=sum(Cash)) %>%
  as_tsibble(index="Month")

forecast::autoplot(tsATM4c)

```



### ATM 4

1. Examine the data.

ATM4 has an extreme outlier which must be rectified. We can use Arima interpolation.

```{r, fig.height=2}
train <- Summarize_ATM(tsATM4, "ATM4")
```
After the interpolation there are still some high values for cash, but we keep them since they are merely the tail end of a range of values. The variance looks smaller toward the end of the series so we will employ a boxcox transformation.

```{r}

tsATM4a <- tsATM4  %>%
  mutate(Cash=ifelse(Cash>10000, NA, Cash))

tsATM4a <- tsATM4a %>%
  model(ARIMA(Cash)) %>%
  interpolate(tsATM4a)


```

```{r}
train <- Summarize_ATM(tsATM4a, "ATM4")
```
We will train the model on 75% of the data.

```{r}
lambda <- tsATM4a %>%
  features(Cash, features = guerrero) %>%
  pull(lambda_guerrero)

tsATM4b <- tsATM4a %>%
 mutate(Cash=box_cox(Cash, lambda))

train <- Summarize_ATM(tsATM4b, "ATM4 with BoxCox")

fit <- Train_ATM(train[[1]])
```
```{r, include=FALSE}
pp <- Examine_Residuals(fit)
```
```{r, fig.height=9}
Print_Residuals(pp)
```
```{r, fig.height=4}
ww <- Check_RMSE(fit, train[[2]], tsATM4b, "ATM2")
plot(ww[[1]])
ww[[3]]
```

3. Forecast using the best model.

ARIMA performs the best - comparable to the lowest RMSE with normal looking residuals centered on zero and looking like white noise.  

The model is an (0-0-0), (2-0-0)(7) model, showing how important seasonality is in the estimate. The model estimates $9,931 for May.  This is lower than last May and April and probably reflects the downward trend in the data since January.

```{r}
fit4 <- tsATM4b %>%
   model(ARIMA(Cash ~ pdq(0,0,0) + PDQ(2,0,0)))  

fc4 <- fit4 %>%
  forecast(h = 31)

rr <- as.vector(fc4$.mean)
rr4 <- forecast::InvBoxCox(rr, lambda)

fc4 %>%
  forecast::autoplot(tsATM4b) +
  geom_line(aes(y = .fitted), col="#D55E00",
            data = augment(fit4)) +
  labs(y="Count", title="ATM4 - ARIMA(0-0-0)(2-0-0)(7)") +
  guides(colour = "none")

report(fit4)
sum(rr4)

```
```{r}


tsATM4c <- tsATM4a %>%
  as.data.frame() %>%
  group_by(Month=yearmonth(xDate)) %>%
  dplyr::summarize(MonthlyCash=sum(Cash)) %>%
  as_tsibble(index="Month")

forecast::autoplot(tsATM4c)

```

### KWH

1. Examine the data.

The data has both a missing value and an outlier.  We use Arima to interpolate both.

```{r}

dfForecast <- as.data.frame(read_excel("D:\\RStudio\\CUNY_624\\Project1\\ForecastLoad.xlsx"))

tsForecast <- dfForecast |>
  rename("Month" = "YYYY-MMM") |>
  mutate(Month=yearmonth(Month)) |>
  as_tsibble(index=Month)
  

summary(dfForecast)
str(tsForecast)
autoplot(tsForecast, KWH)

```

missing data
```{r}

InterpolateBetweenRows <- function(df, column, rowNumber) {

lo = rowNumber-1
hi = rowNumber+1

df1 <- filter(df, row_number() == lo)
df2 <- filter(df, row_number() == hi)
x = (df1[,column] + df2[,column])/2
df[rowNumber, column] = x
return (df)


}

tsForecastTest <- tsForecast %>%
  dplyr::filter(row_number()>120 & row_number()<140)
autoplot(tsForecastTest, KWH)

which(is.na(tsForecast$KWH))

tsForecast2 <- InterpolateBetweenRows(tsForecast, "KWH", 129)

tsForecastTest <- tsForecast2 %>%
  dplyr::filter(row_number()>120 & row_number()<140)
autoplot(tsForecastTest, KWH)

```

Could do average interpolation, or multiply by 7



```{r}

boxplot(tsForecast2$KWH)
min <- tsForecast2 |> dplyr::filter(KWH == 770523)

tsForecastTest2 <- tsForecast2 %>%
  dplyr::filter(CaseSequence>873 & CaseSequence<893)
autoplot(tsForecastTest2, KWH)

tsForecast3 <- tsForecast2 |>
  mutate(KWH = ifelse(CaseSequence==883,7705230,KWH)) 
  
tsForecastTest3 <- tsForecast3 %>%
  dplyr::filter(CaseSequence>873 & CaseSequence<893)
autoplot(tsForecastTest3, KWH)

min2 <- InterpolateBetweenRows(tsForecast, "KWH", 11)
min2[11, "KWH"]

```

```{r}
tsForecast4 <-tsForecast3 %>%
  dplyr::select(KWH)

  stl<- tsForecast4 %>%
  stl(s.window ="periodic")
  forecast::autoplot(stl)


```

Test  Models KWH


```{r}

tsForecast5 <- tsForecast4 |>
  mutate(KWH=log(KWH))
  
train1 <- tsForecast5 %>%
  filter_index(~ "2009 Dec")

test1 <- tsForecast5 %>%
  filter_index("2010 Jan" ~ .)

fit <- train1 %>%
  model(SNAIVE = SNAIVE(), ETS_AAN = ETS(KWH ~ error("A") + trend("A") + season("N")), ETS_AAA = ETS(KWH ~ error("A") + trend("A") + season("A")), ETS_AAdA = ETS(KWH ~ error("A") + trend("Ad", phi=.85) + season("A")), ARIMA=ARIMA(KWH))

resids <- function(model) {
a <- fit[model] %>% gg_tsresiduals + labs(title=model)
return (a)
}
```

```{r, include=FALSE}

z <- list()
plot_list <- list()

for (i in names(fit)) {
  index=1
  
  x <- resids(i)
  p <- eval(substitute(x, list(i=index)))
  plot_list[[i]] <- p 
  
  index=index+1
}
```

```{r, fig.height=9}

grid.arrange(grobs=c(plot_list$SNAIVE[1:3], plot_list$ETS_AAN[1:3], plot_list$ETS_AAA[1:3], plot_list$ETS_AAdA[1:3], plot_list$ARIMA[1:3]), ncol=3)

```

```{r, fig.height=4}

fc <- fit %>%
  forecast(test1)

fc %>% forecast::autoplot(tsForecast5) + labs(title = "KWH") 

fc %>% accuracy(test1)
```
