---
title: "Eric_Hirsch_624_Homework_1"
output: html_document
date: "2023-01-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Exercises 2.1-2.5 and 2.8

1. Use the help function to explore what the series gafa_stock, PBS, vic_elec and pelt represent.
* Use autoplot() to plot some of the series in these data sets.
* What is the time interval of each series?

```{r}

library(tidyverse)
library(tsibbledata)
library(tsibble)
library(forecast)
library(fpp3)
library(fable)

Explore <- function(ts)
{
  x <- deparse(substitute(ts))

  help(x)
  print(autoplot(ts) +
    ggtitle(paste("Timeplot for ", x)))
  
  print(paste("The time interval for", x, ":  "))
  tsibble::interval(ts)
}

```
```{r}

Explore(gafa_stock)
Explore(vic_elec)
Explore(pelt)

head(PBS)
PBS_A01 <- PBS %>% filter(ATC2 == "A01")
Explore(PBS_A01)

```

2. Use filter() to find what days corresponded to the peak closing price for each of the four stocks in gafa_stock.

```{r}

tsGF <- gafa_stock %>%
  group_by(Symbol) %>%
  dplyr::filter(Close == max(Close))

tsGF


```

```{r}

tute1 <- readr::read_csv("tute1.csv")
View(tute1)

mytimeseries <- tute1 %>%
  mutate(Quarter = yearquarter(Quarter)) %>%
  as_tsibble(index = Quarter)

#without facet_grid
g <- mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() +
  ggtitle("No facet_grid")
g

#with facet_grid
g1 <- g +  facet_grid(name ~ ., scales = "free_y") +
  ggtitle("Has facet_grid")
g1

```

Create time plots of the following four time series: Bricks from aus_production, Lynx from pelt, Close from gafa_stock, Demand from vic_elec.

Use ? (or help()) to find out about the data in each series.
For the last plot, modify the axis labels and title.

```{r}



```

