---
title: "Eric_Hirsch_624_Homework_1"
output: html_document
date: "2023-01-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Exercises 2.1-2.5 and 2.8

2.1. Use the help function to explore what the series gafa_stock, PBS, vic_elec and pelt represent.
* Use autoplot() to plot some of the series in these data sets.
* What is the time interval of each series?

```{r}

library(tidyverse)
library(tsibbledata)
library(tsibble)
library(forecast)
library(fpp3)
library(fable)

Explore <- function(ts)
{
  x <- deparse(substitute(ts))

  help(x)
  print(autoplot(ts) +
    ggtitle(paste("Timeplot for ", x)))
  
  print(paste("The time interval for", x, ":  "))
  tsibble::interval(ts)
}

```
```{r}

Explore(gafa_stock)
Explore(vic_elec)
Explore(pelt)

head(PBS)
PBS_A01 <- PBS %>% filter(ATC2 == "A01")
Explore(PBS_A01)

```

2.2. Use filter() to find what days corresponded to the peak closing price for each of the four stocks in gafa_stock.

```{r}

tsGF <- gafa_stock %>%
  group_by(Symbol) %>%
  dplyr::filter(Close == max(Close))

tsGF


```

2.3 Check what happens when you don’t include facet_grid().

```{r}

tute1 <- readr::read_csv("tute1.csv")

mytimeseries <- tute1 %>%
  mutate(Quarter = yearquarter(Quarter)) %>%
  as_tsibble(index = Quarter)

#without facet_grid
g <- mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() +
  ggtitle("No facet_grid")
g

#with facet_grid
g1 <- g +  facet_grid(name ~ ., scales = "free_y") +
  ggtitle("Has facet_grid")
g1

```

2.4 The USgas package contains data on the demand for natural gas in the US.

Install the USgas package.
Create a tsibble from us_total with year as the index and state as the key.
Plot the annual natural gas consumption by state for the New England area (comprising the states of Maine, Vermont, New Hampshire, Massachusetts, Connecticut and Rhode Island).

```{r}
states = c("Maine", "Vermont", "New Hampshire", "Massachusetts", "Connecticut", "Rhode Island")

tsUSGas_NE <- as_tsibble(USgas::us_total, index=year, key=state) %>%
  filter(state==states)

autoplot(tsUSGas_NE)
  

  


```


2.5 Download tourism.xlsx from the book website and read it into R using readxl::read_excel().

Create a tsibble which is identical to the tourism tsibble from the tsibble package.

Find what combination of Region and Purpose had the maximum number of overnight trips on average.

Create a new tsibble which combines the Purposes and Regions, and just has total trips by State.


```{r}

dfTourism <- readxl::read_excel("tourism.xlsx")
head(dfTourism)

head(tsibble::tourism)

tsTourism <- dfTourism %>%
  mutate(Quarter =yearquarter(Quarter)) %>%
  as_tsibble(key = c(Region, State, Purpose), index = Quarter)

dfTourismdf <- as.tibble(tsTourism)

dfTourismMax <- dfTourismdf %>%
  group_by(Region, Purpose) %>%
  summarize(AveTrips = mean(Trips)) %>%
  arrange(desc(AveTrips))

dfTourismMax

tsTourismSimple <- tsTourism %>%
  group_by(State) %>%
  summarize(Trips = sum(Trips)) %>%
    as_tsibble(key = State, index = Quarter)
  

```
8. Monthly Australian retail data is provided in aus_retail. Select one of the time series as follows (but choose your own seed value):

set.seed(12345678)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))
Explore your chosen retail time series using the following functions:

autoplot(), gg_season(), gg_subseries(), gg_lag(),

ACF() %>% autoplot()

Can you spot any seasonality, cyclicity and trend? What do you learn about the series?

```{r}

set.seed(042760)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))

```

```{r}

autoplot(myseries)
gg_season(myseries) 
gg_subseries(myseries)
gg_lag(myseries)

myseries %>% ACF(Turnover) %>% autoplot()

```


