---
title: "Eric_Hirsch_624_Homework_2"
output: html_document
date: "2023-01-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}

library(tidyverse)
library(tsibbledata)
library(tsibble)
library(forecast)
library(fpp3)
library(fable)

```

Exercises 1,2,3,4,5,7,8,9

__*3.1 Consider the GDP information in global_economy. Plot the GDP per capita for each country over time. Which country has the highest GDP per capita? How has this changed over time?*__

```{r}

global_economy_PerCapita <- global_economy %>%
  mutate(GDPperCapita = round(GDP/Population,0)) %>%
  select(GDPperCapita) %>%
  drop_na()

  global_economy_PerCapita %>% autoplot()
  
```

3.2 __*For each of the following series, make a graph of the data. If transforming seems appropriate, do so and describe the effect. United States GDP from global_economy, slaughter of Victorian “Bulls, bullocks and steers” in aus_livestock, Victorian Electricity Demand from vic_elec, Gas production from aus_production.__*


```{r}

ge_US <- global_economy %>%
  mutate(GDPperCapita = round(GDP/Population,0)) %>%
  filter(Country=="United States") %>%
  mutate(logGDPperCapita = log(GDPperCapita))

ge_US %>% autoplot(logGDPperCapita)

```
```{r}


aus_production %>%
  autoplot(Gas)

```
```{r}

ge_US <- global_economy %>%
  mutate(GDPperCapita = round(GDP/Population,0)) %>%
  filter(Country=="United States") 

ge_US %>%
  autoplot(GDPperCapita)

```

