---
title: "Eric_Hirsch_624_Homework_6"
output: html_document
date: "2023-3-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}

library(tidyverse)
library(tsibble)
library(fpp3)
library(EHData)
library(gridExtra)
library(fable)

```

1,2,3,5,6,7,8

__*9.1 Figure 9.32 shows the ACFs for 36 random numbers, 360 random numbers and 1,000 random numbers.*__

__9.1a Explain the differences among these figures. Do they all indicate that the data are white noise?*__

Yes, there is no patterning and no correlations exceed .05.

__*9.1b Why are the critical values at different distances from the mean of zero? Why are the autocorrelations different in each figure when they each refer to white noise?.*__

They are constructed from different sample sizes. The larger the sample, the more "white noise" the sample will be.

__*9.2 A classic example of a non-stationary series are stock prices. Plot the daily closing prices for Amazon stock (contained in gafa_stock), along with the ACF and PACF. Explain how each plot shows that the series is non-stationary and should be differenced.*__

The plot of the time series shows a clear trend upward.  The ACF shows an extreme level of autocorrelation.  The PACF shows the extreme degree to which each observation is correlated with the previous one.

```{r}

tsAmazon <- gafa_stock %>%
  filter(Symbol=="AMZN") 

  autoplot(tsAmazon) + labs(title = "AMZN Closing Price")

  
tsAmazon |> ACF(Close) %>% autoplot() + labs(title="AMZN Closing Price")
tsAmazon |> PACF(Close) %>% autoplot() + labs(title="AMZN Closing Price")

```


__*9.3 For the following series, find an appropriate Box-Cox transformation and order of differencing in order to obtain stationary data.*__

__* 9.3a Turkish GDP from global_economy.*__

The boxcox transformation is not necessary to smooth the variance, but does make the curve more linear.  One difference is all that is necessary to create a stationary series.

```{r}

Get_lambda <- function(ts, column) {
  lambda <- ts %>%
  features(ts[,column], features = guerrero) %>%
  pull(lambda_guerrero)

  return(lambda)

}

tsTurk <- global_economy %>%
  filter(Country=="Turkey")

tsTurk <- tsTurk |>  mutate(GDP=box_cox(GDP, Get_lambda(tsTurk, "GDP")))

  tsTurk %>% autoplot(GDP)+ labs(title="Turkey GDP")
  tsTurk |> ACF(GDP) %>% autoplot() + labs(title="Turkey GDP")
  tsTurk |> PACF(GDP) %>% autoplot() + labs(title="Turkey GDP")
  
  tsTurk |>
  mutate(diff_GDP = difference(GDP)) |>
  features(diff_GDP, ljung_box, lag = 10)

  tsTurk |> ACF(difference(GDP)) |>
  autoplot() + labs(subtitle = "Changes in Turkey GDP")

  tsTurk |> PACF(difference(GDP)) |>
  autoplot() + labs(subtitle = "Changes in Turkey GDP")
  
  tsTurk |>
  mutate(diff_GDP = difference(GDP)) |>
  features(diff_GDP, ljung_box, lag = 10)
  
```

__*9.3b Accommodation takings in the state of Tasmania from aus_accommodation.*__

```{r}

tsAcc <- aus_accommodation %>%
  filter(State=="Tasmania") 

tsAcc <- tsAcc |>  mutate(Takings=box_cox(Takings, Get_lambda(tsAcc, "Takings")))

  tsAcc %>% autoplot(Takings)+ labs(title="Tasmania Takings")
  tsAcc |> ACF(Takings) %>% autoplot() + labs(title="Tasmania Takings")
  tsAcc |> PACF(Takings) %>% autoplot() + labs(title="Tasmania Takings")

  tsAcc |> ACF(difference(Takings)) |>
  autoplot() + labs(subtitle = "Changes in Tasmania Takings")

  tsAcc |> PACF(difference(Takings)) |>
  autoplot() + labs(subtitle = "Changes in Tasmania Takings")
  
  tsAcc |>
  mutate(diff_Takings = difference(Takings)) |>
  features(diff_Takings, ljung_box, lag = 10)
  
  tsAcc |> ACF(difference(difference(Takings, 4), 1)) |>
  autoplot() + labs(subtitle = "Changes*2 in Tasmania Takings")

  tsAcc |> PACF(difference(difference(Takings, 4), 1)) |>
  autoplot() + labs(subtitle = "Changes*2 in Tasmania Takings")
  
  tsAcc |>
  mutate(diff_Takings = difference(difference(Takings, 4), 1)) |>
  features(diff_Takings, ljung_box, lag = 10)

```

__*9.3c Monthly sales from souvenirs.*__

```{r}

tsSouvenirs <- souvenirs %>%
  mutate(Sales=box_cox(Sales, Get_lambda(souvenirs, "Sales")))

  tsSouvenirs %>% autoplot(Sales)+ labs(title="Souvenir Sales")
  tsSouvenirs |> ACF(Sales) %>% autoplot() + labs(title="Souvenir Sales")
  tsSouvenirs |> PACF(Sales) %>% autoplot() + labs(title="Souvenir Sales")

  tsSouvenirs |> ACF(difference(Sales)) |>
  autoplot() + labs(subtitle = "Changes in Souvenir Sales")

  tsSouvenirs |> PACF(difference(Sales)) |>
  autoplot() + labs(subtitle = "Changes in Souvenir Sales")
  
  tsSouvenirs |>
  mutate(diff_Sales = difference(Sales)) |>
  features(diff_Sales, ljung_box, lag = 10)
  
  tsSouvenirs |> ACF(difference(difference(Sales, 12), 1)) |>
  autoplot() + labs(subtitle = "Changes*2 in Souvenir Sales")

  tsSouvenirs |> PACF(difference(difference(Sales, 12), 1)) |>
  autoplot() + labs(subtitle = "Changes*2 in Souvenir Sales")
  
  tsSouvenirs |>
  mutate(diff_Sales = difference(Sales, 12)) |>
  features(diff_Sales, ljung_box, lag = 10)


```

__*9.5 For your retail data (from Exercise 8 in Section 2.10), find the appropriate order of differencing (after transformation if necessary) to obtain stationary data.*__

```{r}

set.seed(042760)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1)) %>%
  fill_gaps() %>%
  dplyr::select(Turnover)

myseries <- myseries %>%
  mutate(Turnover=box_cox(Turnover, Get_lambda(myseries, "Turnover")))

  myseries %>% autoplot(Turnover)+ labs(title="turnover")
  myseries |> ACF(Turnover) %>% autoplot() + labs(title="turnover")
  myseries |> PACF(Turnover) %>% autoplot() + labs(title="turnover")
  
  forecast::ndiffs(myseries$Turnover)

  myseries |> ACF(difference(Turnover)) |>
  autoplot() + labs(subtitle = "Changes in turnover")

  myseries |> PACF(difference(Turnover)) |>
  autoplot() + labs(subtitle = "Changes in turnover")
  
  myseries |>
  mutate(diff_Turnover = difference(Turnover)) |>
  features(diff_Turnover, ljung_box, lag = 10)
  
  myseries |> ACF(difference(difference(Turnover, 6), 1)) |>
  autoplot() + labs(subtitle = "Changes*2 in turnover")

  myseries |> PACF(difference(difference(Turnover, 6), 1)) |>
  autoplot() + labs(subtitle = "Changes*2 in turnover")
  
  myseries |>
  mutate(diff_Turnover = difference(Turnover, 6)) |>
  features(diff_Turnover, ljung_box, lag = 10)


```
__*9.6 Simulate and plot some data from simple ARIMA models.*__

```{r}

y <- numeric(100)
e <- rnorm(100)
for(i in 2:100)
  y[i] <- 0.6*y[i-1] + e[i]
sim <- tsibble(idx = seq_len(100), y = y, index = idx)
```

__*9.6b Produce a time plot for the series. How does the plot change?*__

The plot becomes denser with decreasing levels of phi.

```{r}

autoplot(sim) +labs(title="phi=.6")
y <- numeric(100)
e <- rnorm(100)
for(i in 2:100)
  y[i] <- 0.9*y[i-1] + e[i]
sim <- tsibble(idx = seq_len(100), y = y, index = idx)

autoplot(sim) +labs(title="phi=.9")

y <- numeric(100)
e <- rnorm(100)
for(i in 2:100)
  y[i] <- 0.1*y[i-1] + e[i]
sim <- tsibble(idx = seq_len(100), y = y, index = idx)

autoplot(sim) +labs(title="phi=.1")

```

__*9.6c Write your own code to generate data from an MA(1) model*__

```{r}
y <- numeric(100)
e <- rnorm(100)
for(i in 2:100)
  y[i] <- e[i] + 0.6*e[i-1]
sim2 <- tsibble(idx = seq_len(100), y = y, index = idx)

```

__*9.6d Produce a time plot for the series. How does the plot change?*__

The plot becomes denser with decreasing levels of phi.

```{r}

autoplot(sim2) +labs(title="phi=.6")
y <- numeric(100)
e <- rnorm(100)
for(i in 2:100)
  y[i] <- e[i] + .9*e[i-1]
sim2 <- tsibble(idx = seq_len(100), y = y, index = idx)

autoplot(sim2) +labs(title="phi=.9")

y <- numeric(100)
e <- rnorm(100)
for(i in 2:100)
  y[i] <- e[i] + .1*e[i-1]
sim2 <- tsibble(idx = seq_len(100), y = y, index = idx)

autoplot(sim2) +labs(title="phi=.1")

```

__*9.6e Generate data from an ARMA(1,1) model.__*

```{r}

y <- numeric(100)
e <- rnorm(100)
for(i in 2:100)
  y[i] <- e[i] + .6*e[i-1] + 0.6*y[i-1]
sim3 <- tsibble(idx = seq_len(100), y = y, index = idx)

```

__*9.6f Generate data from an AR(2) model.*__

```{r}

y <- numeric(100)
e <- rnorm(100)
for(i in 3:100)
  y[i] <- e[i] -.8*y[i-1] + .3*y[i-2]
sim4 <- tsibble(idx = seq_len(100), y = y, index = idx)

```

__*9.6g Graph the latter two series and compare them.*__

```{r}

autoplot(sim3) + labs(title="ARMA")
autoplot(sim4) + labs(title="AR(2)")
```

__*9.7 Consider aus_airpassengers, the total number of passengers (in millions) from Australian air carriers for the period 1970-2011.*__

__9.7a Use ARIMA() to find an appropriate ARIMA model. What model was selected. Check that the residuals look like white noise. Plot forecasts for the next 10 periods.*__

```{r}

fit <- aus_airpassengers %>%
  model(ARIMA(Passengers))

report(fit)

fit %>% gg_tsresiduals()

fc <- fit %>%
  forecast(h=10)

fc2 %>% autoplot(aus_airpassengers) + labs(title = "ARIMA 0-2-1")

```

__*9.7b Write the model in terms of the backshift operator.__*

__*9.7c Plot forecasts from an ARIMA(0,1,0) model with drift and compare these to part a.*__

```{r}

fit2 <- aus_airpassengers %>%
  model(ARIMA(Passengers ~ pdq(0,1,0)))

report(fit2)

fit2 %>% gg_tsresiduals()

fc2 <- fit2 %>%
  forecast(h=10)

fc2 %>% autoplot(aus_airpassengers) + labs(title = + labs(title = "ARIMA 0-1-0")) 

```

__*9.7d Plot forecasts from an ARIMA(2,1,2) model with drift and compare these to parts a and c. Remove the constant and see what happens.*__

```{r}
fit2 <- aus_airpassengers %>%
  model(ARIMA(Passengers ~ pdq(2,1,2)))

report(fit2)

fit2 %>% gg_tsresiduals()

fc2 <- fit2 %>%
  forecast(h=10)

fc2 %>% autoplot(aus_airpassengers) + labs(title = + labs(title = "ARIMA 2-1-2")) 


fit3 <- aus_airpassengers %>%
  model(ARIMA(Passengers ~ pdq(2,1,2)))

report(fit3)

fit3 %>% gg_tsresiduals()

fc3 <- fit3 %>%
  forecast(h=10)

fc3 %>% autoplot(aus_airpassengers) + labs(title = + labs(title = "ARIMA 2-1-2, no drift")) 


```


