---
title: "Eric_Hirsch_624_Homework_1"
output: html_document
date: "2023-01-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

1. Use the help function to explore what the series gafa_stock, PBS, vic_elec and pelt represent.
* Use autoplot() to plot some of the series in these data sets.
* What is the time interval of each series?

```{r}
library(tidyverse)
library(tsibbledata)
library(tsibble)
library(forecast)
library(fpp2)
library(fable)

```

```{r}

dfCSD1 <- readr::read_csv("CSD1.csv") %>%
  mutate(DateOfService=as.Date(DateOfService), Month=yearmonth(DateOfService)) %>%
  rename(Count=CountOfServiceID) %>%
  group_by(Month) %>%
  summarize(Count=sum(Count))

tsCSD1 <- as_tsibble(dfCSD1, index=Month)

#TimePlot
autoplot(tsCSD1)

#without facet_grid
g <- mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() +
  ggtitle("No facet_grid")
g

#with facet_grid
g1 <- g +  facet_grid(name ~ ., scales = "free_y") +
  ggtitle("Has facet_grid")
g1



```


```{r}
Explore <- function(ts)
{
  x <- deparse(substitute(ts))

  print("----------------------------------------")
  print(x)
  print(help(ts))
  print(summary(ts))
  print(autoplot(ts) +
    ggtitle(paste("Autoplot for ", x)))
  
  print(paste("Interval for", x, ":  "))
  interval(ts)
}

timePlot <- function(ts, x, y, color) {
  
g <- ts %>%
  pivot_longer(-x) %>%
  ggplot(aes_string(x = pivot, y = y, color = color)) +
  geom_line() +
  ggtitle("No facet_grid") +  
  facet_grid(name ~ ., scales = "free_y") 

return (g)
  
}

```
```{r}

Explore(gafa_stock)
Explore(vic_elec)
Explore(pelt)

  print("----------------------------------------")
  print(paste("Interval for", "PBS", ":  "))
  interval(PBS)

```

2. Use filter() to find what days corresponded to the peak closing price for each of the four stocks in gafa_stock.

```{r}

tsGF <- gafa_stock %>%
  group_by(Symbol) %>%
  dplyr::filter(Close == max(Close))

tsGF


```

```{r}

tute1 <- readr::read_csv("tute1.csv")
View(tute1)

mytimeseries <- tute1 %>%
  mutate(Quarter = yearquarter(Quarter)) %>%
  as_tsibble(index = Quarter)

#without facet_grid
g <- mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() +
  ggtitle("No facet_grid")
g

#with facet_grid
g1 <- g +  facet_grid(name ~ ., scales = "free_y") +
  ggtitle("Has facet_grid")
g1

```

Create time plots of the following four time series: Bricks from aus_production, Lynx from pelt, Close from gafa_stock, Demand from vic_elec.

Use ? (or help()) to find out about the data in each series.
For the last plot, modify the axis labels and title.

```{r}



```

